### 2019-01-16 algorithm.tex
re #33: Use model-consistent price index to calculate real GDP in data for Figure 2. Report on alternative CES calibrations and include data-model figures in algorithm.pdf. Update tables.xlsx with new correlation numbers and newly estimated Table 8.

\documentclass[12pt]{article}
\usepackage{amsmath}
\begin{document}
\subsection*{Normalizations and changes in notation}
I set $L_{nt}=1$ for all country and time. This is just a normalization, as the level of producitivity can pick up the slack.

@@

\section{Calibration}
\subsection{Calibrating the CES model}
To calibrate the CES demand shifters $\nu$, we need data on how final expenditure is split across sectors in each country. We will recover this from the split of revenue, subtracting intermediate expenditure.

@@

@@

The parameter vector $\rho$ and $\sigma$ is estimate using simple OLS. Shocks $u$ are drawn for iid standard normal distribution.
\end{document}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}

\begin{document}
\subsection*{Normalizations and changes in notation}
I set $L_{nt}=1$ for all country and time. This is just a normalization, as the level of producitivity can pick up the slack.

@@

\section{Calibration}
\subsection{Calibrating the CES model*}
To calibrate the CES demand shifters $\nu$, we need data on how final expenditure is split across sectors in each country. We will recover this from the split of revenue, subtracting intermediate expenditure.

@@

However, calibrating nontradable prices this way relies on our estimates of final expenditure shares. These tend to be particularly noisy as they involve a matrix inversion and the subtraction of intermediate expenditure, which, in reality, may not be drive by the same technology matrix across countries. Hence our price estimates and productivity estimates are very noisy. This leads to GDP volatility that is much higher than in the data (Figure 1).

\begin{figure}
\includegraphics[width=\linewidth]{CES2-model-data.pdf}
\caption{GDP volatility in the data and in the model calibrated to $\sigma=2$}
\end{figure}

As an alternative calibration, we use the price index formula for the Cobb-Douglas specification,
\[
P_{nJt} = (P_{0t}PWT_{nt}/P_{n.t})^{1/\nu_{Jt}}
\]
with
\[
P_{n.t} \equiv \prod_{j=1}^{J-1}P_{njt}^{\nu_{jt}}.
\]
The benefit of this formula is that it does not depend on the noisily measured country-specific expenditure shares. This leads to the same calibrated productivity process as under the Cobb-Douglas scenario ($\sigma=1$). This provides a much better fit with the data (Figure 2).

\begin{figure}
\includegraphics[width=\linewidth]{EOS2-model-data.pdf}
\caption{GDP volatility in the data and in the model with $\sigma=2$ but productivity calibrated to Cobb-Douglas case}
\end{figure}

@@

The parameter vector $\rho$ and $\sigma$ is estimated using simple OLS. Shocks $u$ are drawn for iid standard normal distribution.
\end{document}
